---
title: "RISK MANAGEMENT PROJECT"
author: "GROUP3:Neha Sharma, Betty Guo, Hongxi Wan, Nishant Agrawal and Monal Nasa"
date: "4/13/2021"
output: ioslides_presentation
---

```{r, echo=FALSE}
# setwd("C:/Users/mojmc/Desktop/R 654")
library(RTL)
library(tidyr)
library(tidyverse)
library(tidyquant)
library(ggplot2)
library(timetk)
library(PerformanceAnalytics)
library(fOptions)
library(scales)
```


```{r,echo = FALSE}
##FUTURES-------------------------------------------------------------------

fut_fulljul<- RTL::getPrice(feed="CME_NymexFutures_EOD",contract="@CL21N",
                    from="2021-03-23",
                    iuser="morningstar.dscf@ualberta.ca",
              ipassword = "3P#6%bWv7W")
futjul<- fut_fulljul[1,2]
fut.date <- as.Date("2021-03-23")

fut_fullaug<- RTL::getPrice(feed="CME_NymexFutures_EOD",contract="@CL21Q",
                    from="2021-03-23",
                    iuser="morningstar.dscf@ualberta.ca",
              ipassword = "3P#6%bWv7W")
futaug<- fut_fullaug[1,2]

fut<-  futjul*(16/(16+6))+ (6/22)*futaug

fut<- fut  %>% dplyr::transmute(fut=CL21N) %>% as.numeric()
 

#OPTIONS-------------------------------------------------------------------

optjul<-RTL::getPrice(feed="CME_NymexOptions_EOD",
              contract="@LO21NP5700",
              from = "2021-03-23",
            iuser="morningstar.dscf@ualberta.ca",
              ipassword = "3P#6%bWv7W")
colnames(optjul)[grep("LO21NP5700", colnames(optjul))] <-"optpricejul"
Pjul<- optjul %>% dplyr::filter(date=="2021-03-23") %>% 
  dplyr::transmute(Pjul=optpricejul) %>% as.numeric()

optaug<-RTL::getPrice(feed="CME_NymexOptions_EOD",
              contract="@LO21QP5700",
              from = "2021-03-23",
            iuser="morningstar.dscf@ualberta.ca",
              ipassword = "3P#6%bWv7W")
colnames(optaug)[grep("LO21QP5700", colnames(optaug))] <-"optpriceaug"
Paug<- optaug %>% dplyr::filter(date=="2021-03-23") %>% 
  dplyr::transmute(Paug=optpriceaug) %>% as.numeric()

price<-Pjul*(16/26)+ Paug*(10/26)

#Time to expiry calculation---------------------------------------------------
optexpiry<- as.Date("2021-06-30")
T2expiry<- as.numeric(optexpiry-fut.date)/365#in years

```

```{r, echo=FALSE}
# Western Canadian Select----------------------------------------------------
WCWN<- RTL::getPrice(feed="CME_STLCPC_Futures",contract="WCW2021N",
                    from="2021-03-23",
                    iuser="morningstar.dscf@ualberta.ca",
              ipassword = "3P#6%bWv7W")
WCW_23N <- as.numeric(WCWN[1,2])
WCWQ<- RTL::getPrice(feed="CME_STLCPC_Futures",contract="WCW2021Q",
                    from="2021-03-23",
                    iuser="morningstar.dscf@ualberta.ca",
              ipassword = "3P#6%bWv7W")
WCW_23Q <- as.numeric(WCWQ[1,2])
WCW_23 <- (WCW_23N*16/26)+ (WCW_23Q*10/26)


```


```{r,echo=FALSE}
#CC5 contract----------------------------------------------------------------
CC5N<- RTL::getPrice(feed="CME_STLCPC_Futures",contract="CC52021N",
                    from="2021-03-23",
                    iuser="morningstar.dscf@ualberta.ca",
              ipassword = "3P#6%bWv7W")
CC5_23N <- as.numeric(CC5N[1,2])
CC5Q<- RTL::getPrice(feed="CME_STLCPC_Futures",contract="CC52021Q",
                    from="2021-03-23",
                    iuser="morningstar.dscf@ualberta.ca",
              ipassword = "3P#6%bWv7W")
CC5_23Q <- as.numeric(CC5Q[1,2])


CC5_23 <- (CC5_23N*16/26)+ (CC5_23Q*10/26)


```
## OUR NETBACKS 
We have extracted WCW and CC5 and calculated the IV as on 23rd March 2021:

- WTI price for the month of June is $57.57

- THE WCS differential for the month of June is  -10.5576923076923

- THE CC5 differential for the month of June is  -1.38846153846154

- The IV is 41%

```{r,echo=FALSE}
#IV-------------------------------------------------------------------------
r <- 0.0025
#paste("Risk free rate is:", percent(r)) # Set USD IR at 0.25%

fun <-
   function(x) {
    Black76Option(
    TypeFlag = "p",
      FT = fut,
      X = 57,
      Time = T2expiry,
      r = r,
      sigma = x
    )@price - price
  }
implied.vol <- stats::uniroot(fun, c(0.01, 0.8), tol = 0.001)$root
```
## WTI Futures Curve










```{r,message=FALSE, warning=FALSE,echo=FALSE}
#To explain why weighted average prices are chosen: June contract has 16 trading days belonging to july fut and 6 days belonging to aug fut

date = "2021-03-23"
fig.title = paste("Swap Pricing: WTI Futures Curve as of", date)
feeds = dplyr::tibble(
  feed = c(
    "Crb_Futures_Price_Volume_And_Open_Interest",
    "CME_NymexFutures_EOD_continuous"
  ),
  ticker = c("CL", "CL_001_Month")
)
x <-
  RTL::swapInfo(
    date = date,
    feeds = feeds,
    contract = "cmewti",
    exchange = "nymex",
    iuser="morningstar.dscf@ualberta.ca",
              ipassword = "3P#6%bWv7W",
    output = "all"
  )
curve <- x[[1]]
x[[2]]
suppressMessages(library(dplyr, warn.conflict = FALSE, quietly = TRUE))

```



```{r,echo=FALSE}
#STIMULATIONS WTI---------------------------------------------------------------
set.seed(9876)
S0 <-fut
drift <-  r
sigma <- implied.vol
T <-  T2expiry
periods <-  round(252 * T, 0)
simulations <-  10000
dt <-  T / periods


m <- replicate(simulations, 
               RTL::simGBM(S0,drift,sigma,T2M = T,dt))

m <- as_tibble(m) %>% 
  dplyr::mutate(date = seq(from = Sys.Date(), by = "day", length.out = periods)) 
colnames(m) <- c(paste(c("sim"),c(1:simulations),sep = ""),"date")

gbm <- m[nrow(m),] %>% 
  tidyr::pivot_longer(-date,"series","value")

```

## NCF cash flows before the hedging 




```{r,echo=FALSE}
#NCF CALC AND GRAPH----------------------------------------------------------------------
gbm<-gbm %>% mutate(ncfbefore=(gbm$value + WCW_23) - (gbm$value+CC5_23)*.3 - 10) 

gbm %>% ggplot(aes(x = ncfbefore)) + 
geom_histogram(aes(y = stat(count / sum(count))),bins = 100,col="blue",fill="blue")+
  labs(title = "Density Histogram of NCF Terminal Value before hedging", 
       caption = "",
       y = "Density", x = "")+
       geom_vline(xintercept=mean(gbm$ncfbefore), color="red")#+xlim(-20,20 )
```
- NCF under $12 per barrel is 16.65 % and under 0 is 0.05%


## NCF vs WTI Prices


```{r,echo=FALSE,warning=FALSE}
#WTI vs NCF GRAPH-------------------------------------------------------------
ggplot() + 
geom_line(data = gbm, aes(x = gbm$value, y = gbm$ncfbefore), color = "blue")+
  labs(
    title = "NCF vs WTI Prices",
    y = "NCF ($/bbl)",
    x = "WTI ($/bbl)"
  )


```




```{r,echo=FALSE}
## PERCENTAGE-----------------
percentage_values <- subset(gbm$ncfbefore, gbm$ncfbefore < 0)
#length(percentage_values)/10000*100
percentage_values_12 <- subset(gbm$ncfbefore, gbm$ncfbefore < 12)
#length(percentage_values_12)/10000*100


```






##  DIFFERENT HEDGING STRATEGIES 
- Swaps
- Options
- Futures

AND OUR PROPOSED STRATEGY







## 1) Using Swaps at $57.57


```{r,echo=FALSE}
# Hedging through Commodity Swap:
#SWAP NCF CALC----------------------------------------------------------------

m_swap <- as_tibble(m) %>% 
  dplyr::mutate(date = seq(from = Sys.Date(), by = "day", length.out = periods)) 
colnames(m_swap) <- c(paste(c("sim"),c(1:simulations),sep = ""),"date")

gbm_swap <- m_swap[nrow(m_swap),] %>% 
  tidyr::pivot_longer(-date,"series","value")

gbm_swap$value<- 0*(gbm_swap$value +  WCW_23 - (gbm_swap$value+CC5_23)*.3 - 10)+
                1*(fut +  WCW_23 - (gbm_swap$value+CC5_23)*.3 - 10) 

gbm_swap %>% ggplot(aes(x = value)) + 
geom_histogram(aes(y = stat(count / sum(count))),bins = 100,col="blue",fill="blue")+
  labs(title = "Density Histogram of NCF Terminal Value with Hedging with Swap", 
       caption = "",
       subtitle="NCF under $12 using Swaps is : 2.94 %",
       y = "Density", x = "")+
       geom_vline(xintercept=mean(gbm_swap$value), color="red")#+xlim(-20,20 )
```
```{r,echo=FALSE}
# SWAP % BELOW 0 AND 12/barrel CALC-------------------------------------------
percentage_values_swap <- subset(gbm_swap$value, gbm_swap$value < 0)
percentage_values_12_swap <- subset(gbm_swap$value, gbm_swap$value < 12)
#length(percentage_values_12_swap)/10000*100
```



## 2) Using only Put Option 
- Put Option Premium = $0.34/barrel
```{r,echo=FALSE}
# HEDGING WITH PUT @ 36

## PUT PRICE-------------------------------------------------------------------
opt_put36<-RTL::getPrice(feed="CME_NymexOptions_EOD",
              contract="@LO21NP3600",
              from = "2021-03-23",
            iuser="morningstar.dscf@ualberta.ca",
              ipassword = "3P#6%bWv7W")
colnames(opt_put36)[grep("LO21NP3600", colnames(opt_put36))] <-"optprice36"
price_opt_36<- opt_put36 %>% dplyr::filter(date=="2021-03-23") %>% 
  dplyr::transmute(price=optprice36) %>% as.numeric()

  
```


```{r,message=FALSE, warning=FALSE,echo=FALSE}
# PUT NCF AND GRAPH-------------------------------------------------------------------
putreturn <-ifelse((36-gbm$value)>=0,36-gbm$value,0)
gbm<-gbm %>% mutate(ncfhput=(gbm$value + WCW_23) - (gbm$value+CC5_23)*.3 - 10       -price_opt_36*exp(r*T2expiry) + putreturn)

gbm %>% ggplot(aes(x = ncfhput)) + 
geom_histogram(aes(y = stat(count / sum(count))),bins = 100,col="blue",fill="blue")+
  labs(title = "Density Histogram of NCF Terminal Value using long Put at $36 strike", 
       caption = "",subtitle="NCF under $12 using put option is : 17.8 %",
      
       y = "Density", x = "")+
      geom_vline(xintercept=mean(gbm$ncfhput), color="red")+xlim(-10,90 )
suppressMessages(library(dplyr, warn.conflict = FALSE, quietly = TRUE))


```

```{r,echo=FALSE}
#PUT % BELOW 0 AND 12/barrel-----------------------------------------------------
percentage_values_put <- subset(gbm$ncfhput, gbm$ncfhput < 0)
percentage_values_12_put <- subset(gbm$ncfhput, gbm$ncfhput < 12)
length(percentage_values_12_put)/10000*100


```

## 3) Using only Call Option
- Call Option Premium: $0.35/barrel
```{r,message=FALSE, warning=FALSE,echo=FALSE}
# SHORT A CALL @78
opt_call<-RTL::getPrice(feed="CME_NymexOptions_EOD",
              contract="@LO21NC7800",
              from = "2021-03-23",
            iuser="morningstar.dscf@ualberta.ca",
              ipassword = "3P#6%bWv7W")
colnames(opt_call)[grep("LO21NC7800", colnames(opt_call))] <-"optprice"
price_opt_call<- opt_call %>% dplyr::filter(date=="2021-03-23") %>% 
  dplyr::transmute(price=optprice) %>% as.numeric()

callreturn2 <-ifelse((gbm$value-78)>=0,gbm$value-78,0)

gbm<-gbm %>% mutate(ncfcall_Put=((gbm$value + WCW_23) - (gbm$value+CC5_23)*.3 - 10  + +price_opt_call*exp(r*T2expiry)-callreturn2)) 

# CALL GRAPH--------------------------------------------------------------

gbm %>% ggplot(aes(x = ncfcall_Put)) + 
geom_histogram(aes(y = stat(count / sum(count))),bins = 100,col="blue",fill="blue")+
  labs(title = "Density Histogram of NCF Terminal Value using short Call at $78 strike", 
       caption = "",
       subtitle="NCF under $12 using short call at $78  is : 15.42 %",
      
       y = "Density", x = "")+
      geom_vline(xintercept=mean(gbm$ncfcall_Put), color="red")+xlim(-10,90 )
suppressMessages(library(dplyr, warn.conflict = FALSE, quietly = TRUE))
```
```{r}
# CALL PERCENTAGE--------------------------------------------------------
percentage_values_12_c <- subset(gbm$ncfcall_Put, gbm$ncfcall_Put < 12)
length(percentage_values_12_c)/10000*100

```


```{r,message=FALSE, warning=FALSE,echo=FALSE}
# WEIGHING THE COST AND BENIFIT OF ALL THE STRATEGIES, THE STRATEGY WE SUGGEST 
#CONCLUSIVELY IS A MIXED STRATEGY INCLUDING COLLARS, SWAPS AND UNHEDGED PART OF PRODUCTION

#We are looking at the NCF with 48% collars and 32% swaps and 20% unhedged 
#with put strike at $36 and call strike at $78

gbm<-gbm %>% mutate(ncfcall_hedgestrat=
                      (0.48*((gbm$value + WCW_23) - (gbm$value+CC5_23)*.3 - 10 -price_opt_36 + putreturn+price_opt_call-callreturn2))
                       +(0.32*(fut +  WCW_23 - (gbm$value+CC5_23)*.3 - 10))
                    + (0.20*((gbm$value + WCW_23) -(gbm$value+CC5_23)*.3 - 10)))

#  %s-------------------------------------------------------
percentage_values_12_h <- subset(gbm$ncfcall_hedgestrat, gbm$ncfcall_hedgestrat < 12)
length(percentage_values_12_h)/10000*100

```
## SUGGESTED STRATEGY:
- 48% collars (long put @36 and short call @78) and 32% swaps and 20% unhedged.

- NCF under $12 using our hedging strategy is : 0.31 %
```{r,message=FALSE, warning=FALSE,echo=FALSE}
#SUGGESTED STRATEGY GRAPH----------------------------------------------------

gbm %>% ggplot(aes(x = ncfcall_hedgestrat)) + 
geom_histogram(aes(y = stat(count / sum(count))),bins = 100,col="blue",fill="blue")+
  labs(title = "Density Histogram of NCF Terminal Value", 
       caption = "",
       y = "Density", x = "")+geom_vline(xintercept=mean(gbm$ncfcall_hedgestrat), color="red")+xlim(-10,90 )
```




## Case when WCW is  uncertain
- Generally, Uncertain WCW => Higher volatility => Higher Risk => Higher cost for hedging




##

- Before hedging:
```{r,echo=FALSE}
#STIMUL WCW------------------------------------------------------------
set.seed(9876)
S0 <- WCW_23
drift <-  r
sigma <-0.41# implied.vol
T <-  T2expiry
periods <-  round(252 * T, 0)
simulations <-  10000
dt <-  T / periods

m_wcw <- replicate(simulations, 
               RTL::simGBM(S0,drift,sigma,T2M = T,dt))


```


```{r,echo=FALSE, warning=FALSE,echo=FALSE}
#PUTTING THE STIMUL WTI AND WCW IN ONE TABLE-----------------------------------
m <- as_tibble(m) %>% 
  dplyr::mutate(date = seq(from = Sys.Date(), by = "day", length.out = periods)) 
colnames(m) <- c((c(1:simulations)),"date")

m_wcw <- as_tibble(m_wcw) %>% 
  dplyr::mutate(date = seq(from = Sys.Date(), by = "day", length.out = periods)) 
colnames(m_wcw) <- c((c(1:simulations)),"date")


gbm <- m[nrow(m),] %>% 
  tidyr::pivot_longer(-date,"series","wti")

gbm_wcw <- m_wcw[nrow(m_wcw),] %>% 
  tidyr::pivot_longer(-date,"series","wcw")


gbm_wti_wcw <- merge(gbm,gbm_wcw, by = c("series","date"))

# CALC NCF IN CASE OF UNCERTAIN WCW------------------------------------

gbm<-gbm_wti_wcw %>% mutate(ncfbef_wti_wcs=(gbm_wti_wcw$value.x + gbm_wti_wcw$value.y) - (gbm_wti_wcw$value.x+CC5_23)*.3 - 10) 

#PLOTTING NCFFOR UNCERTAIN WCW-------------------------------------------------
gbm %>% ggplot(aes(x = ncfbef_wti_wcs)) + 
geom_histogram(aes(y = stat(count / sum(count))),bins = 100,col="blue",fill="blue")+
  labs(title = "Density Histogram of NCF Terminal Value", 
       caption = "",
       subtitle = "NCF under $12 before the hedge is : 8.16 %",
       y = "Density", x = "")+
       geom_vline(xintercept=mean(gbm$ncfbef_wti_wcs), color="red")#+xlim(-20,20 )
```

```{r,echo=FALSE}

# PERCENTAGE FOR UNCERTAIN WCW--------------------------------------------
# Uncertain WCW> higher volatility> higher risk>higher cost to hedge
percentage_values_12_hw <- subset(gbm$ncfbef_wti_wcs, gbm$ncfbef_wti_wcs < 12)
length(percentage_values_12_hw)/10000*100
```
## 


 - After hedging 


```{r,echo=FALSE}
# NCF IN CASE OF UNCERTAIN WCW (AFTER HEDGE)---------------------------------
gbm<-gbm_wti_wcw %>% mutate(ncfbef_wti_wcs_hedge=(0.48*((gbm_wti_wcw$value.x + gbm_wti_wcw$value.y) - (gbm_wti_wcw$value.x+CC5_23)*.3 - 10-price_opt_36 + putreturn+price_opt_call-callreturn2) +0.2*((gbm_wti_wcw$value.x + gbm_wti_wcw$value.y) - (gbm_wti_wcw$value.x+CC5_23)*.3 - 10)+0.32*((fut + gbm_wti_wcw$value.y) - (gbm_wti_wcw$value.x+CC5_23)*.3 - 10))) 

#PLOTTING NCF - UNCERATIN WCW- (AFTER HEDGE)----------------------------------
gbm %>% ggplot(aes(x = ncfbef_wti_wcs_hedge)) + 
geom_histogram(aes(y = stat(count / sum(count))),bins = 100,col="blue",fill="blue")+
  labs(title = "Density Histogram of NCF Terminal Value after the Hedge", 
       caption = "",
       subtitle="NCF under $12 using hedge is : 0.85 %",
       y = "Density", x = "")+
       geom_vline(xintercept=mean(gbm$ncfbef_wti_wcs_hedge), color="red")#+xlim(-20,20 )
```
```{r,echo=FALSE}
# PERCENTAGE AFTER HEDGE- UNCERTAIN WCW-----------------------------------------
percentage_values_12_hwcs <- subset(gbm$ncfbef_wti_wcs_hedge, gbm$ncfbef_wti_wcs_hedge < 12)
#length(percentage_values_12_hwcs)/10000*100
```
## The MTM gain/loss
- The MTM hedged gain as of 27 April 2021 for our suggested strategy is $1.79 per barrel

```{r,echo=FALSE}
#HEDGE MTM GAIN/LOSS ON 27th April 2021
optexpiry1<- as.Date("2021-06-18")

fut1.date <- as.Date("2021-04-27")
T2expiry_new<- as.numeric(optexpiry1-fut1.date)/365#
c1 <- Black76Option("c",50,78,T2expiry_new,.0025,.15)@price
p1 <- Black76Option("p",50,36,T2expiry_new,.0025,.15)@price

per_hzcc<- 0.48 #% #hedged using costless collar
per_hs<-   0.32 #% #hedged using swaps
MTM<- per_hzcc*((price_opt_call-c1)-(price_opt_36-p1))+ per_hs*(55.57-50)


#view(MTM1)
```

```{r,echo=FALSE}
#CALCULATION OF MTM GAIN/LOSS- DIFFERENT WAY- (FULL NCF ON 23rd Mar- FULL NCF on 27th Apr)
a<- (0.48*((50 + WCW_23) - (50+CC5_23)*.3 - 10 ))+(0.32*(57.57 +  WCW_23 - (50+CC5_23)*.3 - 10)) + (0.2*((50 + WCW_23) -(50+CC5_23)*.3 - 10))
b<-  (0.48*((57.57+ WCW_23) - (57.57+CC5_23)*.3 - 10 +0.01   ))+(0.32*(57.57 +  WCW_23 - (57.57+CC5_23)*.3 - 10))+ (0.2*((57.57+ WCW_23) -(57.57+CC5_23)*.3 - 10))

#paste(b-a,"is our gain per barrel")

```




```{r,echo=FALSE}
# SENSITIVITY ANALYSIS FOR A FULLY HEDGED PROFILE 
#MN: use to explain why didn't go for fully hedged prod: because with fully hedged the avg netback is around $20, and we were able to achieve that same avg netback using mixed strategy. Therefore, less cost same avg.
gbm <- m[nrow(m),] %>% 
  tidyr::pivot_longer(-date,"series","value")

changes_percent <- seq(0, 1, 0.01)
original <- mean(gbm$value)
sensitivity_percent <- data.frame(Collar.Weight=changes_percent, Swap.Weight=(1-changes_percent), Avg_NCF = rep(NA, 101),  Cost = rep(NA,101), Probability = rep(NA,101))

for (i in 1:101){
  NCF <- sensitivity_percent$Collar.Weight[i]*((gbm$value + WCW_23) - (gbm$value+CC5_23)*.3 - 10 -price_opt_36*exp(r*T2expiry) +   putreturn+price_opt_call*exp(r*T2expiry)-callreturn2) +(1-sensitivity_percent$Collar.Weight[i])*(fut +  WCW_23 - (gbm$value+CC5_23)*.3 - 10)
  NCF <- sort(NCF)
  CVaR_0.9 <- NCF[1:1000]
  CVaR_0.1 <- NCF[9001:10000]
  CVaR_0.95 <- NCF[1:500]
  CVaR_0.05 <- NCF[9501:10000]
  sensitivity_percent$Avg_NCF[i] <- mean(NCF)
  
  sensitivity_percent$Cost[i] <- (1.3-0.35)*sensitivity_percent$Collar.Weight[i]
  sensitivity_percent$Probability[i] <- sum(NCF < 12)/10000
}


```

## SENSITIVITY ANALYSIS FOR A FULLY HEDGED PROFILE


```{r, echo=F,results='asis',error=F,warning=F}
library('knitr')
kable(head(sensitivity_percent,5), format = "markdown")
```

```{r}

```












## Thank You